---
title: "Hi-MC: Example Analysis"
author: "Nicholas R. Wheeler, Sandra Smieszek, Dana Crawford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2 - Example Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---
# Example Analysis

### Introduction

This vignette contains an example **HiMC** analysis using demonstration data files.

SNP information is loaded from standard *.map* and *.ped* files, and expected classification outcomes are loaded from an independantly generated *.csv* file.
The **HiMC** package is used to rapidly classify the samples into mitochondrial haplogroups, and the results are checked against the expected outcomes.

Finally, standard R functions and plotting methods are used to visualize and further explore the results.

### Loading Data

First, we load the SNP data from standard *.map* and *.ped* files.

The paths to these files are declared using the `system.file()` function, which retrieves file paths relative to a package installation location.

Defining known local file paths directly as strings would also work.

```{r, message=FALSE, warning=FALSE}
# Call library() to load the package
library("HiMC")

# Get file paths to .map and .ped files
map_file_path <- system.file("extdata", "1000g_MT_HIMC.map", package="HiMC")
ped_file_path <- system.file("extdata", "1000g_MT_HIMC.ped", package="HiMC")

# Use file paths to input the .map and .ped files
snp_data_frame <- HiMC::generate_snp_data(map_file_path,ped_file_path)
```

The expected classification outcomes are accessed and loaded into the session similarly.

```{r, message=FALSE, warning=FALSE}
# Get the file path to the .csv file of expected results
exp_file_path <- system.file("extdata", "1000g_MT_HiMC_classifications.csv", package="HiMC")

# Use the file path to load the expected results as a dataframe
exp_data_frame <- read.csv(exp_file_path)
```

The relative sizes of the dataframes give an idea of how things are going so far.

```{r, message=FALSE, warning=FALSE}
# Retrieve dataframe dimensions
snp_dims <- dim(snp_data_frame)
exp_dims <- dim(exp_data_frame)

# Report dataframe dimensions
cat("Size of SNP dataframe:\n rows:",snp_dims[1],"\n cols:",snp_dims[2],"\n",
  "Size of EXP dataframe:\n rows:",exp_dims[1],"\n cols:",exp_dims[2],sep="")
```

The number of rows (samples) in the two dataframes is a match.

So far, so good!

### HiMC Analysis

Next, we use run the classifier on the SNP dataframe:

```{r, message=FALSE, warning=FALSE}
# Get the HiMC mitochondrial haplogroup classifications
classifications <- HiMC::getClassifications(snp_data_frame)
```

This dataframe of classification results should be identical to the expected results dataframe.

```{r, message=FALSE, warning=FALSE}
# Generate dataframe of comparative bools
comparison <- exp_data_frame==classifications

# Display some comparative bool information
print(head(comparison))
# How many fields are in conflict?
print(length(which(comparison==FALSE)))
```

They are a perfect match, excellent!

### Exploring the Results

Now we can use some standard R functions and plotting methods to further explore and visualize the results.

First let's count the number of unique mitochondrial haplogroups.

```{r, message=FALSE, warning=FALSE}
# Extract haplogroups as a vector
haplogroups <- classifications[,"haplogroup"]

# Display the number of unique haplogroups
print(length(unique(haplogroups)))
```

The samples have been classified into 36 unique mitochondrial haplogroups by **HiMC**!

Next, let's use a barplot to visualize the proportion of samples in each of the 36 haplogroups.

```{r, message=FALSE, warning=FALSE, fig.width=7, fig.height=4}
# Get the descending order of the proportion groups
desc_order <- rev(order(prop.table(table(haplogroups))))

# Make a barplot of the sample proportions in descending order
barplot(prop.table(table(haplogroups))[desc_order],las=2)
```

The most populated haplogroup is 'M', and very few of the samples are unclassified.

Awesome!




